### 2013-12-15 paper.tex
Initial skeleton

\documentclass{article}
\usepackage{natbib}

\title{High-Performance Data Analysis on Janus using Apache Spark}
\author{Nick Vanderweit \\
        Ning Gao \\
        Anitha Ganesha \\
        Michael Kasper}

\begin{document}
\maketitle

\bibliography{paper}
\bibliographystyle{plain}

\end{document}

### 2013-12-16 paper.tex
Some work on introduction

@@
 
+\section{Introduction}
+Over the past decade, there has been considerable interest in high-level
+strategies for evaluating algorithms in parallel over large data sets.
+Google's seminal MapReduce paper \citep{dean-mapreduce} describes such a
+general strategy, which had already been in use at Google at the time of
+publication, for building highly-scalable parallel applications out of small
+serial functions. In general, phrasing data parallelism in terms of
+higher-order functions on parallel data structures has proven fruitful
+for many applications. In this paper, we evaluate Spark, a library
+designed to address some of the shortcomings of MapReduce, on an existing
+supercomputer.
+
+In the MapReduce model, data is broken into \emph{splits} of a given size
+before processing. These are stored on a distributed filesystem as
+key/value pairs. The master node then distributes splits among the remaining
+workers. Each of these workers runs a \emph{Map} on its split, computing for each
+key/value pair $(k_1, v_1)$ another pair $(k_2, v_2)$. Each of these output
+pairs is stored on the filesystem and the master is informed of their
+locations.
+
+The master node proceeds by forwarding these intermediate pairs to the
+\emph{Reduce} workers, which are responsible for combining the results
+for each intermediate key.
+
 \bibliography{paper}

### 2013-12-16 paper.tex
added outline to eval, results, and discussion sections

--- 2013-12-15 paper.tex
@@
 
+\section{Introduction}
+
+\section{Overview}
+
+\section{Evaluation}
+% janus
+    % overview
+%%% DIRECTLY COPY-PASTED TEXT %%%
+University of the University of Colorado's Top-
+31 (June 2010) supercomputer system, Janus
+    % specs (ref)
+%%% DIRECTLY COPY-PASTED TEXT %%%
+The final Janus computational
+system consists of 1,368 nodes, each containing two six-
+core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total
+of 2,736 processors and 16,416 cores with a theoretical
+peak performance of 184 TFLOP/s. Each core has 2
+GB of 1333 MHz DDR3 RAM for a total of 24 GB per
+node and 32 TB of memory in the entire system. [REF]
+% spark
+    % reason for spark (diskless)
+    % installation (default???)
+    % version number (Spark 0.8.0, Scala 2.9.3)
+% purpose of trials (speedup, efficiency, install on Janus)
+    % find ideal configurations
+
+\subsection{Testing}
+% grep
+    % overview
+    % program source
+    % grep complexity
+    % file content
+    % files sizes
+    % node core counts
+    % serial runtime
+    % workers per node
+% page rank
+    % overview
+    % ...
+
+\subsection{Results}
+% small grep test
+    % serial run time (graph)
+    % efficiency (graph)
+    % karp flatt (graph)
+% cores grep test
+    % scale up, workers per node (graph)
+% large grep test
+    % serial run time (graph)
+    % efficiency (graph)
+% pagerank test 1
+% pagerank test 2
+% ...
+
+\section{Discussion}
+% small grep test
+    % ideal balance between speedup and efficiency
+% cores grep test
+    % ideal number of workers per core
+    % close enough not to matter
+    % don't use 6-12
+% large grep test
+    % ideal workers per problem size
+    % follow up test with delay
+    % delay (graph)
+    % take away on IO
+% pagerank test 1
+% pagerank test 2
+% ...
+
+% workings and master comm of spark protocol
+
+\section{Conclusion}
+
 \bibliography{paper}

### 2013-12-16 paper.tex
fixed merge conflicts

@@
 \section{Introduction}
+Over the past decade, there has been considerable interest in high-level
+strategies for evaluating algorithms in parallel over large data sets.
+Google's seminal MapReduce paper \citep{dean-mapreduce} describes such a
+general strategy, which had already been in use at Google at the time of
+publication, for building highly-scalable parallel applications out of small
+serial functions. In general, phrasing data parallelism in terms of
+higher-order functions on parallel data structures has proven fruitful
+for many applications. In this paper, we evaluate Spark, a library
+designed to address some of the shortcomings of MapReduce, on an existing
+supercomputer.
+
+In the MapReduce model, data is broken into \emph{splits} of a given size
+before processing. These are stored on a distributed filesystem as
+key/value pairs. The master node then distributes splits among the remaining
+workers. Each of these workers runs a \emph{Map} on its split, computing for each
+key/value pair $(k_1, v_1)$ another pair $(k_2, v_2)$. Each of these output
+pairs is stored on the filesystem and the master is informed of their
+locations.
+
+The master node proceeds by forwarding these intermediate pairs to the
+\emph{Reduce} workers, which are responsible for combining the results
+for each intermediate key.
 

### 2013-12-16 paper.tex
More MapReduce stuff for introduction

@@
 In the MapReduce model, data is broken into \emph{splits} of a given size
-before processing. These are stored on a distributed filesystem as
-key/value pairs. The master node then distributes splits among the remaining
-workers. Each of these workers runs a \emph{Map} on its split, computing for each
+before processing. These are stored on a distributed filesystem as key/value
+pairs. The master node then distributes splits among the remaining workers.
+Each of these workers runs a \emph{Map} on its split, computing for each
 key/value pair $(k_1, v_1)$ another pair $(k_2, v_2)$. Each of these output
 pairs is stored on the filesystem and the master is informed of their
-locations.
+locations. The master node proceeds by forwarding these intermediate pairs to the
+\emph{Reduce} workers, which are responsible for combining the results for each
+intermediate key.
 
-The master node proceeds by forwarding these intermediate pairs to the
-\emph{Reduce} workers, which are responsible for combining the results
-for each intermediate key.
+A commonly-used example for this process is a distributed word count. First,
+the input file (e.g. a large database dump containing posts on an Internet
+forum) is broken into splits of a given size (say 64MB). The splits store
+$(k, v)$ pairs such that each $v$ is a record (a post). The \emph{Map} task
+is responsible for mapping each record to a set of (word, count) pairs,
+which are written to the filesystem.
+
+The next step is for a \emph{Reduce} worker to sort these records by key
+and then sum together the values for each given key, and write out
+a set of (word, count) records where the keys are now unique.
+
+This simple example has interesting implications for MapReduce in general.
+We can see that \emph{Map} tasks are highly-parallel, but \emph{Reduce}
+introduces a serial bottleneck. In cases like parallel word count, where the
+\emph{Reduce} is associative and commutative (i.e. forms a commutative
+semigroup over the set of values), there is an additional opportunity for
+parallelism. As such, MapReduce provides an additional task, called a
+\emph{Combiner}, which assumes this structure and can be used to distill
+the intermediate $(k, v)$ pairs before a \emph{Reduce}.
 

### 2013-12-16 RPic/R.R
add some pictures generated using R

@@ -1,0 +1,28 @@
+#3. insufficient mem
+memtest=data.frame(memlimit=c('18GB','10GB','6GB','2GB','1GB'),time=c(168L,160L,248L,544L,800L))
+barplot(memtest$time,names.arg=memtest$memlimit,ylim=c(0,700),main="Insufficient Mem Test",xlab="Mem Limit(GB)",ylab="Time(Sec)",col="lightblue")
+
+#1. scalability
+numofnodes=c(2,3,4,5,6)
+time=c(70,73.67,57,52,58.67)
+barplot(time,names.arg=numofnodes,ylim=c(0,80),main="Scalability Test",xlab="Num of Nodes",ylab="Time(Sec)",col=c("lightblue"))
+#2. in memory iterations
+iter=c(1,2,3,4,5,6,7,8,9,10,11)
+nodes12=c(62,17,11,31,23,13,10,13,15,12,20)
+nodes8=c(46,11,8,6,6,8,7,10,8,9,6)
+iter_time=rbind(nodes8,nodes12)
+barplot(as.table(iter_time),names.arg=iter,col=c("red","lightblue"),beside=TRUE,main="Iteration Time",xlab="Num of Iteration",ylab="Time(Sec)", legend = rownames(iter_time),ylim=c(0,70))
+
+#5. data skew
+nodeindex=c(1,2,3,4)
+datainmem=c(403439616,1210318843,1344798720,1492579123)
+barplot(datainmem,names.arg=nodeindex,main="Data Skew",ylab="Bytes In Mem",xlab="Node Index",col=c("lightblue"))
+nodeindex=c(1,2,3,4,5,6,7,8)
+datainmem=c(437059584,551220019,504299520,403439616,706019328,638779392,706019328,504299520)
+
+#*. dataset vertices, links
+datasize=c('0.7G','4.2G','7.0G','9.9G')
+numofver=c(42642,639973,1436185,2453307)
+numofavglink=c(114.57,46.88,34.81,28.53)
+barplot(numofver,names.arg=datasize,col=c("red"),main="Num of Vertices",ylim=c(0,2500000),xlab="Data size",ylab="Num of vertices")
+barplot(numofavglink,names.arg=datasize,col=c("lightblue"),main="Average Num of Links",ylim=c(0,120),xlab="Data size",ylab="Average Number of links")

### 2013-12-17 paper.tex
minor changes to outline
added images directory
Finish introduction (?)

@@
 
+We can also see from this example a significant limitation of MapReduce: each
+intermediate $(k, v)$ set is written to the distributed filesystem.  As such,
+in order for MapReduce to be efficient, the individual tasks must be very
+large.
+
+However, some algorithms are not well-described by a simple map-and-then-reduce
+pattern, and require multiple iterations.  If each iteration is small, the
+algorithm's performance will be bottlenecked by this file I/O. For these
+problems, another approach is needed.
+
+Spark was designed at the UC Berkeley AMPLab to address these concerns using a
+more flexible abstraction for distributed data \citep{zaharia}. Rather than
+assuming that all intermediate data is stored on a filesystem, Spark uses
+\emph{Resilient Distributed Datasets}, or RDDs, to represent data that is
+stored on disk, cached in memory, or possibly has not been computed yet at all
+\citep{zaharia_rdd}.  It is primarily targeted toward Scala, a statically-typed
+functional language for the Java Virtual Machine.
+
 \section{Overview}
@@
 GB of 1333 MHz DDR3 RAM for a total of 24 GB per
-node and 32 TB of memory in the entire system. [REF]
+node and 32 TB of memory in the entire system. (REF)
 % spark
@@
     % serial run time (graph)
-    % efficiency (graph)
 % pagerank test 1

### 2013-12-17 paper.tex
removed image float on grep graphics

@@
 \usepackage{natbib}
+\usepackage{graphicx}
+\usepackage{float}
 
@@
     % overview
+    % describe what spark test where
 %%% DIRECTLY COPY-PASTED TEXT %%%
@@
 \subsection{Results}
-% small grep test
-    % serial run time (graph)
-    % efficiency (graph)
-    % karp flatt (graph)
 % cores grep test
     % scale up, workers per node (graph)
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/workerPerNodeTimes.png}
+        \caption{Grep runtimes with different workers per node over 20GB file}
+        \label{fig:workNodeTime}
+    \end{figure}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/workerPerNodeSpeedup.png}
+        \caption{Grep speedup with different workers per node over 20GB file}
+    \end{figure}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/workerPerNodeEfficiency.png}
+        \caption{Grep efficiency with different workers per node over 20GB file}
+        \label{fig:workNodeEff}
+    \end{figure}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/workerPerNodeKarpFlatt.png}
+        \caption{Grep Karp-Flatt metric with different workers per node over
+                20GB file}
+        \label{fig:workNodeKF}
+    \end{figure}
+        % stays relative constant, serial is the issue
 % large grep test
     % serial run time (graph)
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/bigDataTimes.png}
+        \caption{Average grep runtimes over 100GB file}
+        \label{fig:bigTime}
+        % stays relative constant, serial is the issue
+    \end{figure}
 % pagerank test 1
@@
     % follow up test with delay
-    % delay (graph)
     % take away on IO

### 2013-12-17 RPic/graph-big.r
added R scripts for graphing line-plots

@@ -1,0 +1,42 @@
+library(methods)
+library(ggplot2)
+library(reshape2)
+
+# example input file:
+#    NODES,TIME
+#    50,555.02
+#    50,544.99
+#    40,516.40
+#    40,511.94
+#    30,515.93
+#    30,525.12
+
+# file dim in inches
+width=6
+height=4
+
+# read in raw csv data
+testData <- read.csv("big", header=TRUE)
+
+# get data columns
+nodes <- testData[, "NODES"]
+time <- testData[, "TIME"]
+
+# compute average runtimes
+meanTimes <- tapply(time, nodes, mean)
+
+# add node count column
+meanTimes <- cbind(c(5, 10, 20, 30, 40, 50), meanTimes)
+meanTimes <- data.frame(meanTimes)
+
+# get column headers
+colnames(meanTimes) <- c("NODES", "TIME")
+
+# create graph
+timeGraph <- ggplot(meanTimes, aes(x=NODES, y=TIME)) +
+             geom_line(size=1.5) +
+             ylab("TIME (s)") +
+             scale_y_continuous(limits=c(0, 800))
+
+# write graph to file
+ggsave("big-runtimes.png", w=width, h=height)

### 2013-12-17 RPic/graph.r

@@ -1,0 +1,71 @@
+library(methods)
+library(ggplot2)
+library(reshape2)
+
+# example input file:
+#    NODES,WORKERS,TIME
+#    1,1,111.34
+#    1,1,121.47
+#    1,2,131.66
+#    1,2,121.36
+#    1,6,135.80
+#    1,6,145.08
+#    1,12,144.85
+#    1,12,134.85
+
+# makes graph and writes to file
+# rawData: trial runtimes
+# yLabel: label for y-axis
+# outFile: output file for graph image
+makeGraph <- function(rawData, yLabel, outFile)
+{
+    # image dim in inches
+    width=6
+    height=4
+
+    # compute average, group by nodes+workers
+    avgData <- tapply(rawData, list(nodes, workers), mean)
+
+    # add node count column
+    avgData <- cbind(c(1, 2, 3, 4, 5), avgData)
+    avgData <- data.frame(avgData)
+
+    # add column headers
+    colnames(avgData) <- c("NODES", "1", "2", "3", "6", "12")
+    avgData <- melt(avgData, id=1)
+    colnames(avgData) <- c("NODES", "WORKERS", "YAXIS")
+
+    # create graph
+    timeGraph <- ggplot(avgData, aes(x=NODES, y=YAXIS, color=WORKERS)) +
+                 geom_line(size=1.5) +
+                 ylab(yLabel)
+
+    # save image to paper directory
+    ggsave(outFile, w=width, h=height)
+}
+
+# unix grep serial runtime (sec)
+serialTime <- 1120
+
+# uncached runtime padding
+timePad = 100
+
+# read in raw csv data
+testData <- read.csv("runtime-list", header=TRUE)
+
+# get data columns
+nodes <- testData[, "NODES"]
+workers <- testData[, "WORKERS"]
+time <- testData[, "TIME"] + timePad
+
+# compute metrics
+processes <- 12 * nodes
+speedup <- serialTime / time
+efficiency <- speedup / processes
+karpFlatt <- (1 / speedup - 1 / processes) / (1 - 1 / processes)
+
+# create graphs
+makeGraph(time, "TIME (s)", "runtime.png")
+makeGraph(speedup, "SPEEDUP", "speedup.png")
+makeGraph(efficiency, "EFFICIENCY", "efficiency.png")
+makeGraph(karpFlatt, "KARP-FLATT", "karp_flatt.png")

### 2013-12-17 paper.tex
added rought outline of grep sections

@@
     % describe what spark test where
-%%% DIRECTLY COPY-PASTED TEXT %%%
-University of the University of Colorado's Top-
-31 (June 2010) supercomputer system, Janus
-    % specs (ref)
-%%% DIRECTLY COPY-PASTED TEXT %%%
-The final Janus computational
-system consists of 1,368 nodes, each containing two six-
-core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total
-of 2,736 processors and 16,416 cores with a theoretical
-peak performance of 184 TFLOP/s. Each core has 2
-GB of 1333 MHz DDR3 RAM for a total of 24 GB per
-node and 32 TB of memory in the entire system. (REF)
-% spark
+    % installed and tested on janus
+    % janus specs
+    % spark, scala version
+    % test to determine performance
     % reason for spark (diskless)
@@
     % version number (Spark 0.8.0, Scala 2.9.3)
-% purpose of trials (speedup, efficiency, install on Janus)
     % find ideal configurations
+
+        % we conducted our evaluaion of Spark
+        % on University of Colorado's Janus super computer
+        % Janus consists of 1,368 nodes, each contains
+        % two size core Intel Xeon Westmere-EP chips at 2.8 GHz
+        % total of 12 cores per node
+        % Each core has 2 GB of 1333 MHZ DDR Ram
+        % 24 GB per node (REF)
+
+        % our project used Spark version 0.8.0
+        % along with Scala 2.9.3
+        % accessed files both from Janus's Lustre and ???
+        % aim to assess spark performance on janus
+        % determine optimal configurations
 
@@
     % overview
-    % program source
-    % grep complexity
-    % file content
+        % used example from googles first paper (ref)
+        % searching for a pattern in large text file
+        % return line count with matches
     % files sizes
+        % two different tests, 20GB, 100GB
+        % memory limitation of 1 node 24 gb
     % node core counts
+        % small test ranged from 1 - 5 nodes
+        % large test ranged from 5 - 50 nodes
+        % 10 trials were performed on each config
+        % each test utilized all 12 cores
     % serial runtime
+        % the spark parallel time was compared to serial
+        % serial time was unix grep command
     % workers per node
+        % other configuration setting was workers per node
+        % what is worker? (ref)
+        % compared on small test
+
+        % following exampe of googles first paper on MapReduce
+        % first test performed a parallel grep over large text file
+        % returning line count of found results
+        % two different tests, different files sizes
+        % required to fit entire file in memory
+        % first test, over 1-5 nodes on 20GB file
+        % second test, over 5-50 nodes on 100GB file
+        % each trials utilzed all 12 cores of each node
+        % 10 trials were performed with each configuration
+        % to compute average runtimes
+        % to compute speedup, the unix grep runtime was used a serial
+        % finally, during first test over 20GB file
+        % played with number of workers per core
+        % workers execute application on their portion of distrubted data set
+
 % page rank
@@
     % scale up, workers per node (graph)
+    % runtimes of several combinations of workers per node
+    % runtimes compare with the serial runtime of 20gb file of 1120s
+    % perhaps 2-3 workers has a slight performance increase over 1 worker
+    % clear that 6 and 12 workers are not ideal
+    % additional overhead being the problem
+
+        % first testing obverve scaling on 20GB grep task
+        % also play with workers per node
+        % see scales and ideal config
+        % figRef shows rutimes grep 20GB file from 1-5, 12-60 processes
+        % comparing to the unix grep serial runtime of 1,120 seconds
+        % each line represents a different config of workers per core
+        % clearly see that 6 and 12 were significanlty slower
+        % most likely due to additional communication overhead
+        % while 1-3 workers relatively similar runtimes
+        % maybe be a slight performance increase using 2-3 workers per node
+
     \begin{figure}[H]
@@
     \end{figure}
+
+    % speedup representation (ref)
+    % starts with 12 cores, fare less than linear speedup
+    % begins to level out around 4 nodes
+    % 1-3 workers actually starts to decrease
+
+        % imageRef illustrates this test in terms of speedup
+        % as one nodes consists of 12 cores
+        % the initial speedup shown of 3-4 shows poor scaling
+        % also see speedup begins to level out around 4 nodes
+        % even observe a decrease in 1-3 workers per node at 5 nodes
+
     \begin{figure}[H]
@@
     \end{figure}
+
+    % efficiency graph (ref)
+    % starts are less than 40% efficiency
+    % drops sharply there after
+    % just about 10% efficiency with 5 nodes, 60 cores
+
+        % in graphRef we can see how well the cores are being utilized
+        % in an efficiency graph
+        % as expected from previous graphs, efficiency not great
+        % starts under 40% efficiency with 1 node
+        % end up at just about 10% efficiency with 5 nodes, or 60 cores
+
     \begin{figure}[H]
@@
     \end{figure}
+
+    % karp-flatt graph (ref)
+    % exception of violent changing 12 workers per node
+    % stays relative constant between 0.14 and 0.12
+    % suggests serial is the issue
+
+        % finally we see this data in terms of karp-flatt metric in graphRef
+        % with the exception of the 12 workers per node in pink
+        % we see that the karp-flatt metric remains fairly constaint
+        % values between 0.12 and 0.14
+        % this constant value would suggest that the serial runtime
+        % is what is preventing better scaling
+
     \begin{figure}[H]
@@
         \includegraphics[width=90mm]{images/workerPerNodeKarpFlatt.png}
-        \caption{Grep Karp-Flatt metric with different workers per node over
-                20GB file}
+        \caption{Grep Karp-Flatt with different workers per node over 20GB file}
         \label{fig:workNodeKF}
     \end{figure}
-        % stays relative constant, serial is the issue
+
 % large grep test
     % serial run time (graph)
+    % want to increase problem size to see better scaleup
+    % clearly see no change in runtimes
+    % longer compute time, PageRank, may differ
+
+        % to see if we could see better scaling over a largest file
+        % switch from 20GB to 100GB file
+        % accessed file from lustre drive
+        % as 100GB cannot fit in memory of one node, start with 5
+        % graphRef should the runtimes over this large file
+        % clearly there appears to be no change in runtimes
+        % remains round 500 seconds between 5 nodes and 50
+        % suspecting short compute time of grep expression
+        % switch to a more computational expensive and iterative PageRank
+
     \begin{figure}[H]
@@
         \label{fig:bigTime}
-        % stays relative constant, serial is the issue
     \end{figure}
+
 % pagerank test 1
@@
 \section{Discussion}
-% small grep test
+% cores grep test
     % ideal balance between speedup and efficiency
-% cores grep test
-    % ideal number of workers per core
+    % ideal number of workers per core 2-3
     % close enough not to matter
     % don't use 6-12
+
+        % the results for the grep test over the 20gb file
+        % gave us some insight on how well such a problem scale in spark
+        % for that given task and problem size
+        % 3 nodes appears to be a good balance between speedup and efficiency
+        % we all saw how workers per node effects runtime
+        % while 1-3 workers remains close, clear 6-12 is not ideal
+        % it seems as if one my observer slightly better results
+        % with 2-3 nodes over 1 node, but not by much
+
 % large grep test
-    % ideal workers per problem size
+    % obvious problem with scaling
+    % take away on IO
     % follow up test with delay
-    % take away on IO
+    % possible evidence from previous karp-flatt, serial time is factor
+    % suspect file is being read in serial
+    % grep can process file as load, no change in time
+    % possible problem with configurations
+
+        % in the large grep test over the 100 gb file
+        % suprisingly obvserved no change in performance
+        % perhaps the results of karp-flatt metric in previous test
+        % gives us same insight as to why we are not seeing improvement
+        % we suspect the long load times are are bigger
+        % than the relatively short time needed to graph the loaded information
+        % the long load times may suggest the file is not read in parallel
+
+        % to explore this issue further, we changed our grep program
+        % short pause after processing each line
+        % this would negate the effects of the serial load
+        % see the problem scale
+
 % pagerank test 1
@@
 
-% workings and master comm of spark protocol
+% other problems exhibited
+    % easy to install, hard to configure/profile/understand results
+    % sporadic runtimes
+
+% workers and master communication over spark protocol
 

### 2013-12-17 paper.tex
wrote evaluation overview, grep test overview, and runtime and speedup results

--- 2013-12-17 paper.tex
@@
 \section{Evaluation}
-% janus
-    % overview
-    % describe what spark test where
-%%% DIRECTLY COPY-PASTED TEXT %%%
-University of the University of Colorado's Top-
-31 (June 2010) supercomputer system, Janus
-    % specs (ref)
-%%% DIRECTLY COPY-PASTED TEXT %%%
-The final Janus computational
-system consists of 1,368 nodes, each containing two six-
-core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total
-of 2,736 processors and 16,416 cores with a theoretical
-peak performance of 184 TFLOP/s. Each core has 2
-GB of 1333 MHz DDR3 RAM for a total of 24 GB per
-node and 32 TB of memory in the entire system. (REF)
-% spark
-    % reason for spark (diskless)
-    % installation (default???)
-    % version number (Spark 0.8.0, Scala 2.9.3)
-% purpose of trials (speedup, efficiency, install on Janus)
-    % find ideal configurations
+We conducted our evaluation of the Spark framework on the University of
+Colorado's Janus supercomputer.  Janus consists of 1,368 nodes, each containing
+two six core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total of 12 cores
+per node.  Additionally, each core has 2 GB of 1333 MHz DDR RAM, for a total of
+24 GB per node \cite{JanusPaper}. In our research, we employed Spark version
+0.8.0 along with Scala version 2.9.3. During our testing of the Spark framework,
+we access input files from Janus' **lustre distrubted file system**. % WHAT ELSE
+Our aim was to assess the Spark on Janus and determine optimal configurations
+to achieve the best performance.
 
 \subsection{Testing}
-% grep
-    % overview
-    % program source
-    % grep complexity
-    % file content
-    % files sizes
-    % node core counts
-    % serial runtime
-    % workers per node
+Following the example from Google's first paper on MapReduce, the first test we
+performed was a parallel grep. Here an application would scan through a large
+test file, looking for lines that matched a given regular expression, finally
+returning the total number of matched lines. Two series of test were run on two
+different files sizes. We first wanted to see how Spark scales in an ideal
+setting, where the entire data set fits in memory. ***This way we avoid the
+large performance hit Spark takes writing intermediate results to disk***.
+Given that each node has a total of 24 GB of RAM, we limited our first problem
+size to 20GB. This allowed a single node to contain the entire problem in
+memory, leaving some space for the OS and other supporting processes. We
+excuted this first test with 1 to 5 nodes. The second test we performed was
+run with 5 to 50 nodes, each processing a 100GB file.
+
+Ten trials were performed for each test, allows us to compute average
+runtimes. To compute the speedup, Spark runtimes were compared to the UNIX grep
+commands serial runtime on the same file. Finally, during our first test on the
+20 GB file, we evaluated performance using a different number of workers per
+node.  In Spark, a worker is an executor with it's own JVM which processes it's
+own portion of the distributed data set. By default, a single worker is
+employed on each node. For this test, we wanted to see if this was the ideal
+setting on Janus.
+
 % page rank
@@
 \subsection{Results}
-% cores grep test
-    % scale up, workers per node (graph)
+In the first Grep test, we examined how Spark scales over a 20GB file using 1
+to 5 nodes. Additionally, each trial was executed with a different number
+workers per node. Figure \ref{fig:workNodeTime} shows the average runtimes as
+the number of nodes increases. Each line in this graph represents a different
+number of workers per node. We clearly can see from this graph that 6 and 12
+workers per node yields significantly worse performance, while the results
+produced by 1 to 3 workers is relatively equal.
+
     \begin{figure}[H]
@@
     \end{figure}
+
+Figure \ref{fig:workNodeSpeed} illustrates the results in terms of speedup.
+Here we are comparing the resulting runtimes to the UNIX grep serial runtime of
+1,120 seconds, processing the same 20 GB file. Note that here 1 node represents
+12 cores.  So the initial speedup between 3 and 4 is not as impressive as one
+might expect.  We also see that the speedup begins to level out around 4 nodes,
+and in the case of 1 to 3 workers per node, it even decreases from 4 to 5
+nodes.
+
     \begin{figure}[H]
@@
         \caption{Grep speedup with different workers per node over 20GB file}
+        \label{fig:workNodeSpeed}
     \end{figure}
+
+% in graphRef we can see how well the cores are being utilized
+% in an efficiency graph
+% as expected from previous graphs, efficiency not great
+% starts under 40% efficiency with 1 node
+% end up at just about 10% efficiency with 5 nodes, or 60 cores
+
     \begin{figure}[H]
@@
     \end{figure}
+
+% finally we see this data in terms of karp-flatt metric in graphRef
+% with the exception of the 12 workers per node in pink
+% we see that the karp-flatt metric remains fairly constaint
+% values between 0.12 and 0.14
+% this constant value would suggest that the serial runtime
+% is what is preventing better scaling
+
     \begin{figure}[H]
@@
         \includegraphics[width=90mm]{images/workerPerNodeKarpFlatt.png}
-        \caption{Grep Karp-Flatt metric with different workers per node over
-                20GB file}
+        \caption{Grep Karp-Flatt with different workers per node over 20GB file}
         \label{fig:workNodeKF}
     \end{figure}
-        % stays relative constant, serial is the issue
-% large grep test
-    % serial run time (graph)
+
+% to see if we could see better scaling over a largest file
+% switch from 20GB to 100GB file
+% accessed file from lustre drive
+% as 100GB cannot fit in memory of one node, start with 5
+% graphRef should the runtimes over this large file
+% clearly there appears to be no change in runtimes
+% remains round 500 seconds between 5 nodes and 50
+% suspecting short compute time of grep expression
+% switch to a more computational expensive and iterative PageRank
+
     \begin{figure}[H]
@@
 \section{Discussion}
-% small grep test
-    % ideal balance between speedup and efficiency
-% cores grep test
-    % ideal number of workers per core
-    % close enough not to matter
-    % don't use 6-12
-% large grep test
-    % ideal workers per problem size
-    % follow up test with delay
-    % take away on IO
+% the results for the grep test over the 20gb file
+% gave us some insight on how well such a problem scale in spark
+% for that given task and problem size
+% 3 nodes appears to be a good balance between speedup and efficiency
+% we all saw how workers per node effects runtime
+% while 1-3 workers remains close, clear 6-12 is not ideal
+% it seems as if one my observer slightly better results
+% with 2-3 nodes over 1 node, but not by much
+
+% in the large grep test over the 100 gb file
+% suprisingly obvserved no change in performance
+% perhaps the results of karp-flatt metric in previous test
+% gives us same insight as to why we are not seeing improvement
+% we suspect the long load times are are bigger
+% than the relatively short time needed to graph the loaded information
+% the long load times may suggest the file is not read in parallel
+
 % pagerank test 1

### 2013-12-17 paper.tex
added efficiency, karp-flatt, big-grep results, and grep discussion

@@
 per node.  Additionally, each core has 2 GB of 1333 MHz DDR RAM, for a total of
-24 GB per node \cite{JanusPaper}. In our research, we employed Spark version
+24 GB per node \citep{tufo}. In our research, we employed Spark version
 0.8.0 along with Scala version 2.9.3. During our testing of the Spark framework,
-we access input files from Janus' **lustre distrubted file system**. % WHAT ELSE
+we access input files from Janus' **lustre distributed file system**. % WHAT ELSE
 Our aim was to assess the Spark on Janus and determine optimal configurations
@@
 Following the example from Google's first paper on MapReduce, the first test we
-performed was a parallel grep. Here an application would scan through a large
-test file, looking for lines that matched a given regular expression, finally
-returning the total number of matched lines. Two series of test were run on two
-different files sizes. We first wanted to see how Spark scales in an ideal
-setting, where the entire data set fits in memory. ***This way we avoid the
-large performance hit Spark takes writing intermediate results to disk***.
-Given that each node has a total of 24 GB of RAM, we limited our first problem
-size to 20GB. This allowed a single node to contain the entire problem in
-memory, leaving some space for the OS and other supporting processes. We
-excuted this first test with 1 to 5 nodes. The second test we performed was
-run with 5 to 50 nodes, each processing a 100GB file.
+performed was a parallel grep \citep{dean-mapreduce}. Here an application would
+scan through a large test file, looking for lines that matched a given regular
+expression, finally returning the total number of matched lines. Two series of
+test were run on two different files sizes. We first wanted to see how Spark
+scales in an ideal setting, where the entire data set fits in memory. ***This
+way we avoid the large performance hit Spark takes writing intermediate results
+to disk***.  Given that each node has a total of 24 GB of RAM, we limited our
+first problem size to 20GB. This allowed a single node to contain the entire
+problem in memory, leaving some space for the OS and other supporting
+processes. We executed this first test with 1 to 5 nodes. The second test we
+performed was run with 5 to 50 nodes, each processing a 100GB file.
 
@@
 
-% in graphRef we can see how well the cores are being utilized
-% in an efficiency graph
-% as expected from previous graphs, efficiency not great
-% starts under 40% efficiency with 1 node
-% end up at just about 10% efficiency with 5 nodes, or 60 cores
+In the efficiency graph show in Figure \ref{fig:workNodeEff} we can see just
+how well the cores are being utilized. As expected from the results shown in
+the previous figures, the efficiency is quite poor. Employing a single node
+starts just under 40\% efficiency. By the time we scale the problem to 5 nodes
+(60 cores) we are just above 10\% efficiency.
 
@@
 
-% finally we see this data in terms of karp-flatt metric in graphRef
-% with the exception of the 12 workers per node in pink
-% we see that the karp-flatt metric remains fairly constaint
-% values between 0.12 and 0.14
-% this constant value would suggest that the serial runtime
-% is what is preventing better scaling
+Finally, Figure \ref{fig:workNodeKF} reflects the Karp-Flatt metric for the
+runtime results. With the exception of the 12 workers per node (pink line), we
+we see that the Karp-Flatt metric remains relatively constant, with values
+between 0.12 and 0.14.  This constraint value would suggest that the large serial
+portion of the application is what is preventing use from seeing better
+scaling.
 
@@
 
-% to see if we could see better scaling over a largest file
-% switch from 20GB to 100GB file
-% accessed file from lustre drive
-% as 100GB cannot fit in memory of one node, start with 5
-% graphRef should the runtimes over this large file
-% clearly there appears to be no change in runtimes
-% remains round 500 seconds between 5 nodes and 50
-% suspecting short compute time of grep expression
-% switch to a more computational expensive and iterative PageRank
+To see if we could observe better scaling, we ran the same grep test over a
+100 GB file.  During this test, we also increased the number of nodes used,
+starting with 5 and going to 50 nodes. Given its size, the input text file was
+accessed from the Janus' lustre filesystem. Figure \ref{fig:bigTime} shows the
+runtimes observed over this large problem.  Clearly there appears to be no
+change in runtimes, as the runtimes hover around 500 seconds, regardless of the
+number of nodes used. Suspecting the short computational time of the grep
+search, we began running our tests on the more computationally expensive
+PageRank algorithm.
 
@@
 \section{Discussion}
-% the results for the grep test over the 20gb file
-% gave us some insight on how well such a problem scale in spark
-% for that given task and problem size
-% 3 nodes appears to be a good balance between speedup and efficiency
-% we all saw how workers per node effects runtime
-% while 1-3 workers remains close, clear 6-12 is not ideal
-% it seems as if one my observer slightly better results
-% with 2-3 nodes over 1 node, but not by much
+The results collected from the initial grep test over the 20 GB file gave us
+some insight on how well such a problem scales in the Spark framework.  Given
+that text file of 20 GB, it appears that employing 3 nodes provides a good
+balance between speedup and efficiency.  We also saw how the number of workers
+per node effects the overall runtime. While 1 to 3 workers yield relatively
+similar results, it seems that using 2 to 3 nodes may give a slight performance
+increase. This test would need to be run several more times to achieve more
+more conclusive runtime averages.  It would also be beneficial to increase the
+input file size and node codes, to see if the observed trend continues as the
+problem scales.
 
-% in the large grep test over the 100 gb file
-% suprisingly obvserved no change in performance
-% perhaps the results of karp-flatt metric in previous test
-% gives us same insight as to why we are not seeing improvement
-% we suspect the long load times are are bigger
-% than the relatively short time needed to graph the loaded information
-% the long load times may suggest the file is not read in parallel
+In the grep test over the 100 GB file, surprisingly we observed practically no
+change in runtime performance. Perhaps the Karp-Flatt metric results in the
+previous test gives us some insight as to why this is the case. We suspect the
+long load times on this larger file greatly outweigh any speedup achieved
+through running grep on multiple workers. Given that, we also suspect that
+Spark could be reading this file in serial from the lustre file system. The
+lack of Spark logging and profiling tools makes it difficult to determine
+exactly how long each stage is takes. But further investigation would be needed
+to determine a possible solution or workaround.
 

### 2013-12-17 paper.tex
Minor formatting change to intro
Add some overview stuff

@@
 \usepackage{float}
+\usepackage{listings}
 
@@
 $(k, v)$ pairs such that each $v$ is a record (a post). The \emph{Map} task
-is responsible for mapping each record to a set of (word, count) pairs,
-which are written to the filesystem.
+is responsible for mapping each record to a set of $(\mbox{word}, \mbox{count})$
+pairs, which are written to the filesystem.
 
@@
 \section{Overview}
+One of the key insights of Spark is that reducing the strictness of a
+computation can offer large gains in performance. This lies at the heart
+of the notion of RDDs: it is often more efficient to represent a \emph{recipe}
+for a result than to represent the result itself, especially when the data sets
+are arbitrarily large. Some programmers may be familiar with the related
+notion of a \emph{thunk} in functional programming.
+
+To demonstrate this idea, consider the following code snippet in a
+functional language like ML:
+
+\begin{lstlisting}[language=ML]
+let rdd = ...
+let mapped = map (fun x -> x * 2) rdd
+let filtered = filter (fun x -> x mod 4 == 2) mapped
+let sum = List.fold_left (fun x y -> x + y) 0 filtered
+...
+\end{lstlisting}
+
+In the MapReduce model, there would be a strict filesystem write after each of
+these lines.  Spark does not impose such a restriction, and allows the data to
+persist in memory between operations. In fact, it even goes a step beyond this.
+By default, RDDs are stored lazily as chains of computations that would
+produce an output value. This has several advantages: for one, it is much
+more practical to recover failed nodes with such a recipe in hand. Also,
+storing computations as a lazy data structure permits certain optimizations.
+For instance, the above example contains a \texttt{map} composed with a
+\texttt{filter}. If this computation is expressed as a data structure,
+the Spark environment can dynamically optimize it to:
+
+\begin{lstlisting}[language=ML]
+let rdd = ...
+let filtered = filter (fun x -> (x * 2) mod 4 == 2) mapped
+let sum = List.fold_left (fun x y -> x + y) 0 filtered
+...
+\end{lstlisting}
+
+This kind of optimization is known as \emph{deforestation}, the elimination
+of intermediate data structures, and is used to greatly improve data
+locality \citep{wadler}.
 

### 2013-12-18 paper.tex
More overview

@@
 locality \citep{wadler}.
+
+In general, this way of identifying data with its lineage is a very powerful
+way to perform \emph{batch computations} on data, which means applying the same
+transformation over many elements of a distributed data structure. Like
+MapReduce, Spark is not well-suited for applications where each parallel
+tasks requires communication with other threads.
+
+Spark provides APIs for Java, Python, and Scala, although the Scala API seems
+to be the most common. This is likely because of Spark's heavy use of
+higher-order functions on data structures, which is a staple of
+functional programming.
+
+\emph{Bagel} is an implementation of the Google Pregel graph processing framework
+for Spark, which works on a distributed set of $(k, v)$ pairs, where $k$ is
+a vertex ID and $v$ is a vertex with its state. This allows algorithms like
+PageRank to be written simply, and we use this to evaluate Spark's performance
+on iterative algorithms.
+
+Spark also has a Hive-compatible query system called Shark. Though it is beyond
+the scope of this project, Shark allows efficient interactive data querying
+using an SQL-like syntax.
 

### 2013-12-18 paper.tex
fixed spelling errors in grep text

@@
 \section{Evaluation}
-We conducted our evaluation of the Spark framework on the University of
+We conducted our evaluation of the Spark system on the University of
 Colorado's Janus supercomputer. Janus consists of 1,368 nodes, each containing
@@
 24 GB per node \citep{tufo}. In our research, we employed Spark version
-0.8.0 along with Scala version 2.9.3. During our testing of the Spark framework,
-we access input files from Janus' **lustre distributed file system**. % WHAT ELSE
-Our aim was to assess the Spark on Janus and determine optimal configurations
-to achieve the best performance.
+0.8.0 along with Scala version 2.9.3. During our testing of the Spark system,
+we access input files from Janus' lustre distributed file system. Our aim was
+to assess the Spark runs on Janus and determine optimal configurations to
+achieve the best performance.
 
@@
 expression, finally returning the total number of matched lines. Two series of
-test were run on two different files sizes. We first wanted to see how Spark
-scales in an ideal setting, where the entire data set fits in memory. ***This
-way we avoid the large performance hit Spark takes writing intermediate results
-to disk***.  Given that each node has a total of 24 GB of RAM, we limited our
-first problem size to 20GB. This allowed a single node to contain the entire
-problem in memory, leaving some space for the OS and other supporting
-processes. We executed this first test with 1 to 5 nodes. The second test we
-performed was run with 5 to 50 nodes, each processing a 100GB file.
+tests were run on two different files sizes. During these tests, we wanted to
+see how Spark scales in an ideal setting, where the entire data set fit in
+memory. This way we avoid the large performance hit Spark takes writing
+intermediate results to disk. Given that each node has a total of 24 GB of RAM,
+we limited our first problem size to 20GB. This allowed a single node to
+contain the entire problem in memory, leaving sufficient memory space for the
+OS and other supporting processes. We executed this first test with 1 to 5
+nodes. The second test we performed was run with 5 to 50 nodes, each processing
+a 100GB file.
 
@@
 runtimes. To compute the speedup, Spark runtimes were compared to the UNIX grep
-commands serial runtime on the same file. Finally, during our first test on the
+command's serial runtime on the same file. Finally, during our first test on the
 20 GB file, we evaluated performance using a different number of workers per
@@
 employed on each node. For this test, we wanted to see if this was the ideal
-setting on Janus.
+setting for running Spark on Janus.
 
@@
 \subsection{Results}
-In the first Grep test, we examined how Spark scales over a 20GB file using 1
+In the first grep test, we examined how Spark scales over a 20GB file using 1
 to 5 nodes. Additionally, each trial was executed with a different number
@@
 Finally, Figure \ref{fig:workNodeKF} reflects the Karp-Flatt metric for the
-runtime results. With the exception of the 12 workers per node (pink line), we
-we see that the Karp-Flatt metric remains relatively constant, with values
-between 0.12 and 0.14.  This constraint value would suggest that the large serial
-portion of the application is what is preventing use from seeing better
-scaling.
+runtime results. With the exception of the test using 12 workers per node (pink
+line), we we see that the Karp-Flatt metric remains relatively constant, with
+values between 0.12 and 0.14. This constant value would suggest that the
+large serial portion of the application is what is preventing use from seeing
+better scaling.
 
@@
 100 GB file. During this test, we also increased the number of nodes used,
-starting with 5 and going to 50 nodes. Given its size, the input text file was
-accessed from the Janus' lustre filesystem. Figure \ref{fig:bigTime} shows the
+starting with 5 and going to 50 nodes. Figure \ref{fig:bigTime} shows the
 runtimes observed over this large problem.  Clearly there appears to be no
@@
 number of nodes used. Suspecting the short computational time of the grep
-search, we began running our tests on the more computationally expensive
-PageRank algorithm.
+search was negligible as compared to the file load time, we began running our
+tests on the more computationally expensive PageRank algorithm.
 
@@
 The results collected from the initial grep test over the 20 GB file gave us
-some insight on how well such a problem scales in the Spark framework.  Given
-that text file of 20 GB, it appears that employing 3 nodes provides a good
-balance between speedup and efficiency.  We also saw how the number of workers
-per node effects the overall runtime. While 1 to 3 workers yield relatively
-similar results, it seems that using 2 to 3 nodes may give a slight performance
+some insight on how well such a problem scales in the Spark system. Given a
+text file of 20 GB, it appears that employing 3 nodes provides a good balance
+between speedup and efficiency. We also saw how the number of workers per node
+effects the overall runtime. While 1 to 3 workers yield relatively similar
+results, it seems that using 2 to 3 nodes may give a slight performance
 increase. This test would need to be run several more times to achieve more
 more conclusive runtime averages. It would also be beneficial to increase the
-input file size and node codes, to see if the observed trend continues as the
-problem scales.
+input file size and node counts, to see if the observed trend continues as the
+problem is scaled up.
 
 In the grep test over the 100 GB file, surprisingly we observed practically no
-change in runtime performance. Perhaps the Karp-Flatt metric results in the
+change in runtime performance. Perhaps the Karp-Flatt metric results seen in the
 previous test gives us some insight as to why this is the case. We suspect the
@@
 through running grep on multiple workers. Given that, we also suspect that
-Spark could be reading this file in serial from the lustre file system. The
+Spark could be reading this file serially from the lustre file system. The
 lack of Spark logging and profiling tools makes it difficult to determine
-exactly how long each stage is takes. But further investigation would be needed
+exactly how long each stage is taking. Further investigation would be needed
 to determine a possible solution or workaround.

### 2013-12-18 paper.tex
s/Janus/JANUS/ and conclusion blurb

@@
 
-\title{High-Performance Data Analysis on Janus using Apache Spark}
+\title{High-Performance Data Analysis on JANUS using Apache Spark}
 \author{Nick Vanderweit \\
@@
 We conducted our evaluation of the Spark system on the University of
-Colorado's Janus supercomputer. Janus consists of 1,368 nodes, each containing
+Colorado's JANUS supercomputer. JANUS consists of 1,368 nodes, each containing
 two six core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total of 12 cores
@@
 0.8.0 along with Scala version 2.9.3. During our testing of the Spark system,
-we access input files from Janus' lustre distributed file system. Our aim was
-to assess the Spark runs on Janus and determine optimal configurations to
+we access input files from JANUS' lustre distributed file system. Our aim was
+to assess the Spark runs on JANUS and determine optimal configurations to
 achieve the best performance.
@@
 employed on each node. For this test, we wanted to see if this was the ideal
-setting for running Spark on Janus.
+setting for running Spark on JANUS.
 
@@
 \section{Conclusion}
+We found that, while Spark offers an appealing abstraction layer for
+data-parallel programming, achieving reasonable scalability even on trivial
+problems is still very difficult. Without adequate profiling tools, we found
+ourselves puzzled by the lackluster results of our experiments. With more time,
+we could probably use the Spark logs to better understand where our serial
+bottlenecks are and increase efficiency. It may very well still be worth
+exploring how Spark could be used on JANUS, but it may require rethinking
+how we are using the network and filesystems for this task.
 

### 2013-12-18 paper.tex
More conclusion

@@
 data-parallel programming, achieving reasonable scalability even on trivial
-problems is still very difficult. Without adequate profiling tools, we found
-ourselves puzzled by the lackluster results of our experiments. With more time,
-we could probably use the Spark logs to better understand where our serial
-bottlenecks are and increase efficiency. It may very well still be worth
-exploring how Spark could be used on JANUS, but it may require rethinking
-how we are using the network and filesystems for this task.
+problems is still very difficult. Any abstraction for parallelism would
+seemingly run into this problem: when speedups are not as expected, it is often
+much harder to track down the source of the problem.  The lack of adequate
+profiling tools made the lackluster results of our experiments all the more
+puzzling. With more time, we could probably use the Spark logs to better
+understand where our serial bottlenecks are and increase efficiency. It may
+very well still be worth exploring how Spark could be used on JANUS, but it may
+require rethinking how we are using the network and filesystems for this task.
 

### 2013-12-18 RPic/plot.R
add some pics

@@ -1,0 +1,11 @@
+nodes=c('04','06','08','10','12','14','16','04','06','08','10','12','14','16','04','06','08','10','12','14','16')
+dataset=c('05m','05m','05m','05m','05m','05m','05m','30m','30m','30m','30m','30m','30m','30m','50m','50m','50m','50m','50m','50m','50m')
+speedup=c(1,2,3,4,5,6,7,288.755,211.515,760.85,145.675,608.245,188.255,208.47,718.12,854.34,507.305,444.59,1,2,3)
+avgData=tapply(speedup,list(nodes,dataset),mean)
+avgData=cbind(c(4,6,8,10,12,14,16),avgData)
+avgData=data.frame(avgData)
+colnames(avgData)=c("NODES","05m","30m","50m")
+avgData=melt(avgData,id=1)
+colnames(avgData)=c("NODES","DATA","SPEEDUP")
+ggplot(avgData,aes(x=NODES,y=SPEEDUP,color=DATA))+geom_line(size=1.5)+ylab("Speedup")
+

### 2013-12-18 paper.tex
pagerank analysis

@@
 
+Now we run five tests for pagerank program.\\
+
+We have tested the scalability of Spark system on distributed Grep program and we still want to see whether for iterative algorithms, it’s also scalable by looking at the time as the number of nodes increases.\\
+
+We test Bagel Pagerank on Dataset of five million links, 5 million links, 30 million links and 50 million links and on number of nodes from 4 to 16 with increment of two.
+As shown in Figure \ref{fig:scalability}, for scalability test, the larger the dataset is, the better the scalability is because the ratio of computation to communication is large.
+The reason we ran on a small dataset is a akka failure with a message of “Remote akka client shutdown”.
+We tried to solve the error by modifying the message size and timeout for the akka client but the error still pops up.\\
+
+\begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/in-mem.png}
+        \caption{In-Memory RDD for iterative algorithms}
+        \label{fig:in-memory}
+    \end{figure}
+
+The main motivation to design Spark is Resilient Distributed Dataset, which resides in memory and stays in memory over the iterations, while Hadoop incurs lots of disk IO because it materializes data like the map output in disk before the data is shuffled.
+This test is to check whether the time for later iterations is less compared with that in the first iteration.\\
+
+We test Simple Pagerank on Dataset with 1.4 million pages and 50 million links in total on 8 nodes and 12 nodes.
+At the same time, we make sure the data will fit in memory.
+After that, we set the number of iterations to be 11 and during the first iteration, the program needs to load the data from the disk and then process the data.
+As shown in Figure \ref{fig:in-memory} for both the 8-node cluster and 12-node cluster, the time spent on the first iteration is much higher than all the latter iterations.
+One thing to note is that the time for 12-node cluster is always higher than 8-node cluster because the dataset is small and the performance gain from distributing the work is less than the communication overhead incurred with more nodes.
+In sum, in-memory dataset does improve the performance for iterative algorithms.\\
+
+Spark stores the whole dataset in memory in avoidance of disk IO but when the memory is sufficient, there’s no way to avoid disk IO.
+Spark has its algorithm to swap the least used data to disk and keeps the most used in memory.
+This test is to find whether the performance degrades linearly as the memory amount decreases.\\
+
+\begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/insufficient_memory.png}
+        \caption{Insufficient memory test}
+        \label{fig:insuff-mem}
+    \end{figure}
+
+We test Bagel Pagerank on Dataset with 64,000 pages and 30 million links in total on four nodes in Janus.
+As we limit the memory size from 1gb to 18gb in each node, the time is recorded.
+As shown in the Figure \ref{fig:insuff-mem}, the time stays stable when the memory limit is over 6gb since the program won’t use disk for RDD data storage.
+Once the memory limit is less than the memory required, the performance drops until it fails on one-node cluster, where the java heap exceeds the limit.
+Despite the performance degradation, it’s quite linear.
+Besides, the dataset size is 4.2gb and it consumes around 8.8gb memory per node, which is 35gb memory in total.
+one of the reasons is that Spark system designs the RDD to be immutable and every time a RDD is transformed, a new RDD is created, thus taking up lots of memory.
+The other reason is that Spark replicates the RDD on the go for loss of data.\\
+
+\begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/bagelsimple.png}
+        \caption{Bagel VS. Simple PageRank}
+        \label{fig:bagelsimple}
+    \end{figure}
+
+Google has Pregel for distributed graph algorithms and Spark also has its own version of Pregel called Bagel for graph processing.
+This test is to check whether the bagel version of pagerank outperforms simple version in terms of time.\\
+
+We ran both Bagel version and Simple version of PageRank on number of nodes from 4 to 16 with increment 2 and Dataset with 64,000 pages and 30 million links in total with sufficient memory.
+In Figure \ref{fig:bagelsimple} regardless of time, we can see that bagel version outperforms simple version a little bit.
+The main reason is that for smaller dataset, the overhead of managing vertices, edges and messages are so huge that the benefits of asynchronous background messages that Spark might offer is not that obvious.
+We couldn’t test on big dataset because of akka failure.
+However, it’s worth noting that programming is easier with general concepts of vertices, edges and messages to represent the graph.\\
+
+\begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/skew.png}
+        \caption{Data Skew Test}
+        \label{fig:dataskew}
+    \end{figure}
+
+At last, we will focus on potential data skew problem, which exists in systems like Hadoop. Since with data skew, the strangler could slow down the whole program considerably.\\
+
+We ran Bagel version of PageRank on four nodes and on dataset of 30 million links.
+As shown in Figure \ref{fig:dataskew}, the memory used is quite balanced across the nodes.
+We checked the data size read from the disk and found that all the nodes are reading the exact same amount of data from the disk.
+So the little imbalance came from the RDDs created later, which is partitioned by a certain key.
+\\
+
 % pagerank test 1
@@
 to determine a possible solution or workaround.
+
+To sum up, Spark, with its in memory data store, performs great on iterative algorithms and degrade linearly with insufficient memory.
+With partitioner, the dataset is uniformly distributed across the nodes in avoidance of possible stranglers.
+For scalability test, the time increasing with the number of nodes is due to the overhead of communication for small dataset.
+we believe that solving the akka error to enable the program to run on big dataset like 25gb will show great scalability for number of nodes under ten.
+
+The biggest problem we had in the project is the akka failure, which occurs when we ran a big dataset on a small number of nodes.
+This prevented us from increasing the ratio of computing to communication for scalability test.
+Second, since there is the time limit for a cluster of nodes we allocate, we have to stop the test in the middle and reallocate nodes, which shows inconsistent time metric.
+It might be that the nodes we reallocate have different network topology with congested network in Janus.
+Third, although we ran each test with three repetitions, we saw significant difference in time with exact the same configuration.
+We simply discard the one that is very unlikely compared with the other two.
 

### 2013-12-18 paper.tex
pixel

@@
 
-We have tested the scalability of Spark system on distributed Grep program and we still want to see whether for iterative algorithms, it’s also scalable by looking at the time as the number of nodes increases.\\
+We have tested the scalability of Spark system on distributed Grep program and we still want to see whether for iterative algorithms, it's also scalable by looking at the time as the number of nodes increases.\\
 

### 2013-12-18 paper.tex
Some stylistic changes

@@
 \usepackage{listings}
+\usepackage{url}
 
@@
 change in runtimes, as the runtimes hover around 500 seconds, regardless of the
-number of nodes used. Suspecting the short computational time of the grep
-search was negligible as compared to the file load time, we began running our
-tests on the more computationally expensive PageRank algorithm.
+number of nodes used.
 
@@
 
-Now we run five tests for pagerank program.\\
+Suspecting the short computational time of the grep search was negligible as
+compared to the file load time, we began running our tests on the more
+computationally expensive PageRank algorithm. We ran five tests in total,
+measuring the amount of wall clock time each took to compute.
 
-We have tested the scalability of Spark system on distributed Grep program and we still want to see whether for iterative algorithms, it's also scalable by looking at the time as the number of nodes increases.\\
-
-We test Bagel Pagerank on Dataset of five million links, 5 million links, 30 million links and 50 million links and on number of nodes from 4 to 16 with increment of two.
-As shown in Figure \ref{fig:scalability}, for scalability test, the larger the dataset is, the better the scalability is because the ratio of computation to communication is large.
-The reason we ran on a small dataset is a akka failure with a message of “Remote akka client shutdown”.
-We tried to solve the error by modifying the message size and timeout for the akka client but the error still pops up.\\
+We tested a Bagel implementation of PageRank on several of 5 million links, 30
+million links, and 50 million links. We ran these computations on allocations
+ranging from 4 to 16 nodes, incrementing by two.  As shown in Figure
+\ref{fig:scalability}, for the scalability test, the larger the dataset is, the
+better the scalability is because the ratio of computation to communication is
+large. We had to run on a small dataset because of a technical problem with our
+Spark installation.\footnote{
+\url{https://groups.google.com/forum/#!topic/spark-users/-nsoF7oonOM}}
 

### 2013-12-18 paper.tex
Style fixes

@@
 
-The main motivation to design Spark is Resilient Distributed Dataset, which resides in memory and stays in memory over the iterations, while Hadoop incurs lots of disk IO because it materializes data like the map output in disk before the data is shuffled.
-This test is to check whether the time for later iterations is less compared with that in the first iteration.\\
+The main technical insight behind Spark is the Resilient Distributed Dataset,
+which resides in memory and stays in memory over the iterations, while Hadoop
+incurs I/O penalties because it materializes intermediate data (like the map
+output) on disk before the data is shuffled. As such, we ran a test to see
+how total wall clock time changed from one run to the next.
 
-We test Simple Pagerank on Dataset with 1.4 million pages and 50 million links in total on 8 nodes and 12 nodes.
-At the same time, we make sure the data will fit in memory.
-After that, we set the number of iterations to be 11 and during the first iteration, the program needs to load the data from the disk and then process the data.
-As shown in Figure \ref{fig:in-memory} for both the 8-node cluster and 12-node cluster, the time spent on the first iteration is much higher than all the latter iterations.
-One thing to note is that the time for 12-node cluster is always higher than 8-node cluster because the dataset is small and the performance gain from distributing the work is less than the communication overhead incurred with more nodes.
-In sum, in-memory dataset does improve the performance for iterative algorithms.\\
+We tested a simple PageRank implementation on a dataset with 1.4 million pages
+and 50 million links in total on 8 nodes and 12 nodes, ensuring that there
+is enough heap space for all of the data to fit into memory. We set
+the number of iterations to be 11. During the first iteration, the program
+needs to load the data from the disk and then process the data. As shown in
+Figure \ref{fig:in-memory} for both the 8-node cluster and 12-node cluster, the
+time spent on the first iteration is much higher than all the latter
+iterations. Note that the time for 12-node cluster is always
+higher than 8-node cluster because the dataset is small and the performance
+gain from distributing the work is less than the communication overhead
+incurred with more nodes. In sum, in-memory dataset does improve the
+performance for iterative algorithms.
 
-Spark stores the whole dataset in memory in avoidance of disk IO but when the memory is sufficient, there’s no way to avoid disk IO.
-Spark has its algorithm to swap the least used data to disk and keeps the most used in memory.
-This test is to find whether the performance degrades linearly as the memory amount decreases.\\
+Spark often stores the whole dataset in memory in order to avoid disk I/O.
+However, when there is not sufficient memory to do this, disk I/O is
+unavoidable.  As such, Spark has an algorithm algorithm to swap the least-used
+data to disk and keeps the most-used in memory. This test is to find how
+the performance degrades as we restrict the amount of heap space.
 

### 2013-12-18 RPic/R.R
dpi correction

@@ -1,24 +1,30 @@
 #3. insufficient mem
+png("~/insufficient_memory.png",width=6,height=4,units="in",res=400)
 memtest=data.frame(memlimit=c('18GB','10GB','6GB','2GB','1GB'),time=c(168L,160L,248L,544L,800L))
 barplot(memtest$time,names.arg=memtest$memlimit,ylim=c(0,700),main="Insufficient Mem Test",xlab="Mem Limit(GB)",ylab="Time(Sec)",col="lightblue")
+dev.off()
 
 #1. scalability
 numofnodes=c(2,3,4,5,6)
 time=c(70,73.67,57,52,58.67)
 barplot(time,names.arg=numofnodes,ylim=c(0,80),main="Scalability Test",xlab="Num of Nodes",ylab="Time(Sec)",col=c("lightblue"))
 #2. in memory iterations
+png("~/in-mem.png",width=6,height=4,units="in",res=400)
 iter=c(1,2,3,4,5,6,7,8,9,10,11)
 nodes12=c(62,17,11,31,23,13,10,13,15,12,20)
 nodes8=c(46,11,8,6,6,8,7,10,8,9,6)
 iter_time=rbind(nodes8,nodes12)
 barplot(as.table(iter_time),names.arg=iter,col=c("red","lightblue"),beside=TRUE,main="Iteration Time",xlab="Num of Iteration",ylab="Time(Sec)", legend = rownames(iter_time),ylim=c(0,70))
+dev.off()
 
 #5. data skew
+png("~/skew.png",width=6,height=4,units="in",res=400)
 nodeindex=c(1,2,3,4)
 datainmem=c(403439616,1210318843,1344798720,1492579123)
 barplot(datainmem,names.arg=nodeindex,main="Data Skew",ylab="Bytes In Mem",xlab="Node Index",col=c("lightblue"))
 nodeindex=c(1,2,3,4,5,6,7,8)
 datainmem=c(437059584,551220019,504299520,403439616,706019328,638779392,706019328,504299520)
+dev.off()
 
 #*. dataset vertices, links
 datasize=c('0.7G','4.2G','7.0G','9.9G')

### 2013-12-18 paper.tex
dpi correction
revised PageRank discussion section

--- 2013-12-18 paper.tex
@@
 
-We test Bagel Pagerank on Dataset with 64,000 pages and 30 million links in total on four nodes in Janus.
+We test Bagel Pagerank on Dataset with 64,000 pages and 30 million links in total on four nodes in JANUS.
 As we limit the memory size from 1gb to 18gb in each node, the time is recorded.
@@
 
-To sum up, Spark, with its in memory data store, performs great on iterative algorithms and degrade linearly with insufficient memory.
-With partitioner, the dataset is uniformly distributed across the nodes in avoidance of possible stranglers.
-For scalability test, the time increasing with the number of nodes is due to the overhead of communication for small dataset.
-we believe that solving the akka error to enable the program to run on big dataset like 25gb will show great scalability for number of nodes under ten.
+To sum up, Spark, with its in memory data store, performs great on iterative
+algorithms and degrade linearly with insufficient memory. Using a partitioner,
+the dataset is uniformly distributed across the nodes to avoidance of possible
+stragglers. For the scalability test, the time increase with the number of nodes
+is due to the communication overhead for the small dataset. We believe that
+solving the akka error would enable the program to run on larger datasets, and
+show great scalability nodes counts of 10 and under.
 
-The biggest problem we had in the project is the akka failure, which occurs when we ran a big dataset on a small number of nodes.
-This prevented us from increasing the ratio of computing to communication for scalability test.
-Second, since there is the time limit for a cluster of nodes we allocate, we have to stop the test in the middle and reallocate nodes, which shows inconsistent time metric.
-It might be that the nodes we reallocate have different network topology with congested network in Janus.
-Third, although we ran each test with three repetitions, we saw significant difference in time with exact the same configuration.
-We simply discard the one that is very unlikely compared with the other two.
+The biggest problem we had throughout these tests was the akka failure. This
+occurs when we ran large dataset on a small number of nodes. This prevented us
+from increasing the ratio of computing to communication for scalability test.
+Secondly, due to the time limit for a cluster of allocated nodes on JANUS, we
+have to stop in the middle of testing and reallocate nodes. This often
+reflected in inconsistent results. It might be that the nodes we reallocate
+have different network topology with network congestion. Finally, although we
+ran three trials for each test, we saw a significant difference in time,
+despite running with the same exact configuration. To help address this issue,
+we simply discarded any result that appeared to be an outlier as compared to
+the other two results.
 

### 2013-12-18 paper.tex

--- 2013-12-18 paper.tex
@@
 
+We have tested the scalability of Spark system on distributed Grep program and we still want to see whether for iterative algorithms, it's also scalable by looking at the time as the number of nodes increases.\\
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=90mm]{images/scalability.png}
+        \caption{Scalability Test for Pagerank}
+        \label{fig:bscalability}
+    \end{figure}
+
 We tested a Bagel implementation of PageRank on several of 5 million links, 30

### 2013-12-18 paper.tex
Style!

@@
 20 GB file, we evaluated performance using a different number of workers per
-node. In Spark, a worker is an executor with it's own JVM which processes it's
+node. In Spark, a worker is an executor with its own JVM which processes its
 own portion of the distributed data set. By default, a single worker is
@@
 
-We have tested the scalability of Spark system on distributed Grep program and we still want to see whether for iterative algorithms, it's also scalable by looking at the time as the number of nodes increases.\\
-
     \begin{figure}[H]
@@
 
-We test Bagel Pagerank on Dataset with 64,000 pages and 30 million links in total on four nodes in JANUS.
-As we limit the memory size from 1gb to 18gb in each node, the time is recorded.
-As shown in the Figure \ref{fig:insuff-mem}, the time stays stable when the memory limit is over 6gb since the program won’t use disk for RDD data storage.
-Once the memory limit is less than the memory required, the performance drops until it fails on one-node cluster, where the java heap exceeds the limit.
-Despite the performance degradation, it’s quite linear.
-Besides, the dataset size is 4.2gb and it consumes around 8.8gb memory per node, which is 35gb memory in total.
-one of the reasons is that Spark system designs the RDD to be immutable and every time a RDD is transformed, a new RDD is created, thus taking up lots of memory.
-The other reason is that Spark replicates the RDD on the go for loss of data.\\
+For comparison, we tested a Bagel implementation of PageRank on a dataset with
+64,000 pages and 30 million links in total on four nodes in Janus. As we limit
+the memory size from 1GB to 18GB in each node, the time is recorded. As shown
+in the Figure \ref{fig:insuff-mem}, the time stays stable when the memory limit
+is over 6GB since the program won’t use disk for RDD data storage. Once the
+memory limit is less than the memory required, the performance drops until it
+fails on one-node cluster, where the Java heap allocation is exceeded. Despite
+the performance degradation, it's quite linear. Besides, the dataset size is
+4.2GB and it consumes around 8.8GB memory per node, which is 35GB of memory in
+total. One of the reasons is that RDDs are immutable and every time a RDD is
+transformed, a new RDD is created, thus taking up lots of memory. The other
+reason is that Spark replicates the RDD on the go for loss of data.
 

### 2013-12-18 paper.tex
fix on ref
Estilo!

@@
         \caption{Scalability Test for Pagerank}
-        \label{fig:bscalability}
+        \label{fig:scalability}
     \end{figure}
@@
 However, when there is not sufficient memory to do this, disk I/O is
-unavoidable.  As such, Spark has an algorithm algorithm to swap the least-used
+unavoidable.  As such, Spark has an algorithm to swap the least-used
 data to disk and keeps the most-used in memory. This test is to find how
@@
 
-Google has Pregel for distributed graph algorithms and Spark also has its own version of Pregel called Bagel for graph processing.
-This test is to check whether the bagel version of pagerank outperforms simple version in terms of time.\\
-
-We ran both Bagel version and Simple version of PageRank on number of nodes from 4 to 16 with increment 2 and Dataset with 64,000 pages and 30 million links in total with sufficient memory.
-In Figure \ref{fig:bagelsimple} regardless of time, we can see that bagel version outperforms simple version a little bit.
-The main reason is that for smaller dataset, the overhead of managing vertices, edges and messages are so huge that the benefits of asynchronous background messages that Spark might offer is not that obvious.
-We couldn’t test on big dataset because of akka failure.
-However, it’s worth noting that programming is easier with general concepts of vertices, edges and messages to represent the graph.\\
+In order to measure the relative performance of the Bagel and simple
+implementations, we ran both versions of PageRank on
+number of nodes from 4 to 16 with increment 2 and a dataset with 64,000 pages and
+30 million links in total with sufficient memory. In Figure
+\ref{fig:bagelsimple} regardless of time, we can see that bagel version
+outperforms simple version a little bit. The main reason is that for a smaller
+dataset, the overhead of managing vertices, edges and messages are so huge that
+the benefits of asynchronous background messages that Spark might offer is not
+that obvious. We couldn’t test on big dataset because of the aforementioned
+technical failure. However, it's worth noting that programming is easier with
+general concepts of vertices, edges and messages to represent the graph.
 
@@
 
-At last, we will focus on potential data skew problem, which exists in systems like Hadoop. Since with data skew, the strangler could slow down the whole program considerably.\\
+Finally, we will focus on the potential data skew problem, which exists in systems
+like Hadoop. Since with data skew, a single straggler could slow down the whole
+program considerably.
 
-We ran Bagel version of PageRank on four nodes and on dataset of 30 million links.
-As shown in Figure \ref{fig:dataskew}, the memory used is quite balanced across the nodes.
-We checked the data size read from the disk and found that all the nodes are reading the exact same amount of data from the disk.
-So the little imbalance came from the RDDs created later, which is partitioned by a certain key.
-\\
+We ran our Bagel version of PageRank on four nodes and on dataset of 30 million
+links. As shown in Figure \ref{fig:dataskew}, the memory used is quite balanced
+across the nodes. We checked the data size read from the disk and found that
+all the nodes are reading the exact same amount of data from the disk. So the
+little imbalance came from the RDDs created later, which is partitioned by a
+certain key.
 

### 2013-12-18 paper.tex
revised later pagerank result discussion

--- 2013-12-18 paper.tex
@@
 
-At last, we will focus on potential data skew problem, which exists in systems like Hadoop. Since with data skew, the strangler could slow down the whole program considerably.\\
+Finally, we will focus on potential data skew problem, which exists in systems
+like Hadoop. Since with data skew, a straggling worker could slow down the whole
+program considerably.\\
 
-We ran Bagel version of PageRank on four nodes and on dataset of 30 million links.
-As shown in Figure \ref{fig:dataskew}, the memory used is quite balanced across the nodes.
-We checked the data size read from the disk and found that all the nodes are reading the exact same amount of data from the disk.
-So the little imbalance came from the RDDs created later, which is partitioned by a certain key.
-\\
+We ran the Bagel version of PageRank on four nodes with a dataset of 30 million
+links. As shown in Figure \ref{fig:dataskew}, the memory used is quite balanced
+across the nodes. We used the Spark monitoring logs to confirm that the number
+of bytes read from disk by each node were exactly the same. So the slight skew
+we observed was created later by the RDDs, which are partitioned by a certain
+key. \\
 

### 2013-12-18 paper.tex
adding pagerank overview

--- 2013-12-18 paper.tex
@@
     % ...
+
+\subsubsection{PageRank}   
+Since Spark system is an in-memory cluster computing system, the power of such a system should be exploited by applications that runs several iterations on the same data.
+
+As an example of iterative algorithms the \emph{Page Rank} algorithm was chosen.
+The original pagerank algorithm was developed by Google websearch engine which enabled them to rank the webpages in their search engine.
+
+\emph{Data set}:
+The source of the dataset is DBPedia, a crowd sourced community that extracts structured information from wikipedia and makes it available on the web.
+\citep{[??]} There are different formats of wikipedia datasets available online.
+For our experiments we chose the wikipedia pagelink dataset \citep{[??]} , this contains
+internal links between the DBPedia instances extracted from the internal links of the wikipedia articles making it apt for the any datamining related analysis including Pagerank algorithm.
+Therefore this dataset was chosen for the experiment.
+The overall size of this dataset is 25GB.
+The format of the data is \{ source url \} \{ some configuration \} \{ dest url \}
+For the pagerank algorithm, only the source url and the dest url is needed.Therefore the dataset was was preprocessed to filter out the only the necessary information.
+
+If the entire dataset is assumed to be represented by a huge graph, then here each DBPedia instance maps to a wikipedia article or a page which inturn could be mapped to a vertex of the graph.
+And the links between the instances could be mapped to
+edges of the graph.
+
+\begin{figure}
+\centering
+  \includegraphics[width=80mm]{images/numVertices.png}
+  \caption{The vertices count over various dataset size}
+  \label{fig:numvertices}
+\end{figure}
+
+\begin{figure} 
+  \centering
+  \includegraphics[width=80mm]{images/numLinks.png}
+  \caption{The average edge count per vertex for various dataset size}
+  \label{fig:numlinks}
+\end{figure}
+
+The figure \ref{fig:numvertices} shows the variation in the no of vertices count for different dataset size which from ~40,000 nodes and the figure \ref{fig:numlinks} shows  the average number of links or edges per vertex for a smaller dataset is much higher as compared to the larger dataset.
+The smaller dataset are the first few DBPedia instances of the larger dataset.
+The first few pages probably have more links compared the pages in the end of the dataset and hence the average is linearly decreasing.
+
+\emph{Algorithm:}
+The basic idea of the page rank algorithm is to assign scores to the pages based on the following two aspects of the page.
+	\begin{enumerate}
+		\item {\centering pages with higher incoming links get high rank.}
+		 \item {\centering pages with links from a high rank page also gets high rank.}
+	\end{enumerate}
+
+The pseudocode of the algorithm is as follows:
+\begin{enumerate}
+	\item {\centering Assign every page a rank of 1.}
+	\item {\centering On each iteration have page p contribute to rankp/ |neighbours|p to its neighbors.}
+	\item {\centering Set each page ranks to \(0.15 + 0.85  *  contributions\)}.
+	\end{enumerate}
+
+The scala implementation of this algorithm is as shown below:
+\begin{lstlisting}[language=ML]
+vallinks =// RDD of (url, neighbors) pairs
+varranks =// RDD of (url, rank) pairs
+for(i<- 1 to ITERATIONS) 
+{
+  valcontribs=links.join(ranks).flatMap{
+  case (url, (links, rank)) => links.
+  map( dest=> (dest, rank/links.size))
+  }
+  ranks =contribs.reduceByKey(_ + _).
+  mapValues(0.15 + 0.85 * _)
+}
+ranks.saveAsTextFile(...) 
+\end{lstlisting}
+
+\emph{Bagel}:
+The Bagel programming model was developed by spark.This is the implementation of the graph processing framework developed by google called the \emph{Pregel} system[??].
+Bagel is still in the development stage where in more features are being added to the basic graph computation framework.
+
+Any graph algorithm runs for several iterations and each iteration is called a Superstep.The graph is represented as a distributed dataset of $(K,V)$ pairs.
+Here, the keys are vertex IDs, the values are vertex and associated states of the vertices.Every vertex in the graph will be associated with a user specified function.In each superstep, the algorithm runs the function associated with the vertex using the messages recieved from the previous superstep.
+The output is a new vertex state and a list of all outgoing messages which is then sent to all the other dependent vertices which is then used as input for the next iteration.
+
+For the pagerank algorithm, as previously explained the vertices are pages and the edges are links.The list of messages generated at every iteration is the contribution of the pagerank for all its outgoing links.
+
+Bagel provides APIs for vertices, edges and messages which could be extended to customize for any algorithm.This enables the user to write any compilcated code within few lines of code.
+
+Further, bagel also provides combiner utility where in a reduction operation is applied on the outgoing message which reduces the communication overhead.The aggregator utility helps in reducing the message across all nodes for all vertices of the graph after every superstep.
+
+\emph{Hash Partitioner}:The algorithm makes use of a hash partitioner which is a Controlled data partitioning utility that reduces communication cost.Also, it keeps the information of how data was partitioned across jobs for further reduction of communication cost.
+
+\emph{Kyro serialization}: For the bagel program a special form of serialization technique was used, this  serializes the object sent over the network that will be cached in serialized form.
+The kryo serialization is much faster than the default java serialization used in the spark system as it the communication cost.
 

### 2013-12-18 paper.tex
Typos!
revised first part of page rank description

@@
 \subsubsection{PageRank}   
-Since Spark system is an in-memory cluster computing system, the power of such a system should be exploited by applications that runs several iterations on the same data.
-
-As an example of iterative algorithms the \emph{Page Rank} algorithm was chosen.
-The original pagerank algorithm was developed by Google websearch engine which enabled them to rank the webpages in their search engine.
+Since Spark is an in-memory cluster computing system, the power of such
+a system should be exploited by applications that runs several iterations on
+the same data.  One such iterative algorithm we choose to employ was
+\emph{PageRank}.  The original PageRank algorithm was developed by Google which
+enabled them to rank the webpages in their search engine.
 
 \emph{Data set}:
-The source of the dataset is DBPedia, a crowd sourced community that extracts structured information from wikipedia and makes it available on the web.
-\citep{[??]} There are different formats of wikipedia datasets available online.
-For our experiments we chose the wikipedia pagelink dataset \citep{[??]} , this contains
-internal links between the DBPedia instances extracted from the internal links of the wikipedia articles making it apt for the any datamining related analysis including Pagerank algorithm.
-Therefore this dataset was chosen for the experiment.
-The overall size of this dataset is 25GB.
-The format of the data is \{ source url \} \{ some configuration \} \{ dest url \}
-For the pagerank algorithm, only the source url and the dest url is needed.Therefore the dataset was was preprocessed to filter out the only the necessary information.
+We obtained the data used for these tests from DBPedia, a crowd sourced
+community that extracts structured information from Wikipedia and makes it
+available on the web. \citep{[??]} There are different formats of Wikipedia
+datasets available online. For our experiments we chose the Wikipedia pagelink
+dataset \citep{[??]} , this contains internal links between the DBPedia
+instances extracted from the internal links of the Wikipedia articles making it
+apt for the any datamining related analysis including PageRank algorithm.
+Therefore this dataset was chosen for the experiment.  The overall size of this
+dataset is 25GB.  The format of the data is of the form: \\
 
-If the entire dataset is assumed to be represented by a huge graph, then here each DBPedia instance maps to a wikipedia article or a page which inturn could be mapped to a vertex of the graph.
-And the links between the instances could be mapped to
-edges of the graph.
+\centering
+\{ source url \} \{ some configuration \} \{ dest url \} \\
+
+For the PageRank algorithm, only the source url and the dest url is
+needed. Therefore the dataset was was preprocessed to filter out only the
+necessary information.  If the entire dataset is assumed to be represented by a
+huge graph, then here each DBPedia instance maps to a Wikipedia article or a
+page which inturn could be mapped to a vertex of the graph. The links
+between the instances could then be mapped to edges of the graph.
 
@@
 
-For the pagerank algorithm, as previously explained the vertices are pages and the edges are links.The list of messages generated at every iteration is the contribution of the pagerank for all its outgoing links.
+For the PageRank algorithm, as previously explained the vertices are pages and the edges are links.The list of messages generated at every iteration is the contribution of the PageRank for all its outgoing links.
 
@@
         \includegraphics[width=90mm]{images/scalability.png}
-        \caption{Scalability Test for Pagerank}
+        \caption{Scalability Test for PageRank}
         \label{fig:scalability}
@@
 
-% pagerank test 1
-% pagerank test 2
+% PageRank test 1
+% PageRank test 2
 % ...

### 2013-12-18 paper.tex
more revisions to PageRank description

@@
 expression, finally returning the total number of matched lines. Two series of
-tests were run on two different files sizes. During these tests, we wanted to
+tests were run on two different file sizes. During these tests, we wanted to
 see how Spark scales in an ideal setting, where the entire data set fit in
@@
 Therefore this dataset was chosen for the experiment.  The overall size of this
-dataset is 25GB.  The format of the data is of the form: \\
+dataset is 25GB.  The format of the data is of the form:
 
-\centering
+\begin{center}
 \{ source url \} \{ some configuration \} \{ dest url \} \\
+\end{center}
 
-For the PageRank algorithm, only the source url and the dest url is
-needed. Therefore the dataset was was preprocessed to filter out only the
-necessary information.  If the entire dataset is assumed to be represented by a
-huge graph, then here each DBPedia instance maps to a Wikipedia article or a
-page which inturn could be mapped to a vertex of the graph. The links
-between the instances could then be mapped to edges of the graph.
+For the PageRank algorithm, only the source url and the dest url is needed.
+Therefore the dataset was was preprocessed to filter out only the necessary
+information.  If the entire dataset is assumed to be represented by a huge
+graph, then here each DBPedia instance maps to a Wikipedia article or a page
+which inturn could be mapped to a vertex of the graph. The links between the
+instances could then be mapped to edges of the graph.
 
-\begin{figure}
+\begin{figure}[H]
 \centering
@@
 
-\begin{figure} 
+\begin{figure}[H]
   \centering
@@
 
-The figure \ref{fig:numvertices} shows the variation in the no of vertices count for different dataset size which from ~40,000 nodes and the figure \ref{fig:numlinks} shows  the average number of links or edges per vertex for a smaller dataset is much higher as compared to the larger dataset.
-The smaller dataset are the first few DBPedia instances of the larger dataset.
-The first few pages probably have more links compared the pages in the end of the dataset and hence the average is linearly decreasing.
+The figure \ref{fig:numvertices} shows the variation in the vertex count for
+different dataset sizes which range from approximately 40,000 nodes and the
+figure \ref{fig:numlinks} shows the average number of links or edges per vertex
+for a smaller dataset is much higher as compared to the larger dataset. The
+smaller dataset are the first few DBPedia instances of the larger dataset. The
+first few pages probably have more links compared the pages in the end of the
+dataset and hence the average is linearly decreasing.
 
 \emph{Algorithm:}
-The basic idea of the page rank algorithm is to assign scores to the pages based on the following two aspects of the page.
+The basic idea of the page rank algorithm is to assign scores to the pages
+based on the following two aspects of the page.
+
 \begin{enumerate}
     \item {\centering pages with higher incoming links get high rank.}
-		 \item {\centering pages with links from a high rank page also gets high rank.}
+    \item {\centering pages with links from a high rank page also gets high
+        rank.}
 \end{enumerate}
@@
     \item {\centering Assign every page a rank of 1.}
-	\item {\centering On each iteration have page p contribute to rankp/ |neighbours|p to its neighbors.}
+    \item {\centering On each iteration have page p contribute to $rankp/
+        |neighbours|p$ to its neighbors.}
     \item {\centering Set each page ranks to \(0.15 + 0.85  *  contributions\)}.
@@
 the benefits of asynchronous background messages that Spark might offer is not
-that obvious. We couldn’t test on big dataset because of the aforementioned
+that obvious. We couldn't test on a big dataset because of the aforementioned
 technical failure. However, it's worth noting that programming is easier with

### 2013-12-18 paper.tex
finished revising PageRank overview

@@
 pairs is stored on the filesystem and the master is informed of their
-locations. The master node proceeds by forwarding these intermediate pairs to the
-\emph{Reduce} workers, which are responsible for combining the results for each
-intermediate key.
+locations. The master node proceeds by forwarding these intermediate pairs to
+the \emph{Reduce} workers, which are responsible for combining the results for
+each intermediate key.
 
@@
 
-For the PageRank algorithm, only the source url and the dest url is needed.
-Therefore the dataset was was preprocessed to filter out only the necessary
-information.  If the entire dataset is assumed to be represented by a huge
-graph, then here each DBPedia instance maps to a Wikipedia article or a page
-which inturn could be mapped to a vertex of the graph. The links between the
-instances could then be mapped to edges of the graph.
+For the PageRank algorithm, only the \emph{source url} and the \emph{dest url}
+is needed.  Therefore the dataset was was preprocessed to filter out only the
+necessary information.  If the entire dataset is assumed to be represented by a
+huge graph, then here each DBPedia instance maps to a Wikipedia article or a
+page which in turn could be mapped to a vertex of the graph. The links between
+the instances could then be mapped to edges of the graph.
 
@@
 
-The scala implementation of this algorithm is as shown below:
+The Scala implementation of this algorithm is as shown below:
 \begin{lstlisting}[language=ML]
@@
 \emph{Bagel}:
-The Bagel programming model was developed by spark.This is the implementation of the graph processing framework developed by google called the \emph{Pregel} system[??].
-Bagel is still in the development stage where in more features are being added to the basic graph computation framework.
+The Bagel programming model was developed by Spark. This is the implementation
+of the graph processing framework developed by Google called the \emph{Pregel}
+system[??]. Bagel is still in the development stage where in more features are
+being added to the basic graph computation framework.
 
-Any graph algorithm runs for several iterations and each iteration is called a Superstep.The graph is represented as a distributed dataset of $(K,V)$ pairs.
-Here, the keys are vertex IDs, the values are vertex and associated states of the vertices.Every vertex in the graph will be associated with a user specified function.In each superstep, the algorithm runs the function associated with the vertex using the messages recieved from the previous superstep.
-The output is a new vertex state and a list of all outgoing messages which is then sent to all the other dependent vertices which is then used as input for the next iteration.
+Any graph algorithm runs for several iterations and each iteration is called a
+Superstep. The graph is represented as a distributed dataset of $(K,V)$ pairs.
+Here, the keys are vertex IDs, the values are vertex and associated states of
+the vertices. Every vertex in the graph will be associated with a user specified
+function. In each superstep, the algorithm runs the function associated with the
+vertex using the messages received from the previous superstep. The output is a
+new vertex state and a list of all outgoing messages which is then sent to all
+the other dependent vertices which is then used as input for the next
+iteration.
 
-For the PageRank algorithm, as previously explained the vertices are pages and the edges are links.The list of messages generated at every iteration is the contribution of the PageRank for all its outgoing links.
+For the PageRank algorithm, as previously explained the vertices are pages and
+the edges are links. The list of messages generated at every iteration is the
+contribution of the PageRank for all its outgoing links.
 
-Bagel provides APIs for vertices, edges and messages which could be extended to customize for any algorithm.This enables the user to write any compilcated code within few lines of code.
+Bagel provides APIs for vertices, edges and messages which could be customized
+for any algorithm. This enables the user to write any complicated code with
+only few lines of code.
 
-Further, bagel also provides combiner utility where in a reduction operation is applied on the outgoing message which reduces the communication overhead.The aggregator utility helps in reducing the message across all nodes for all vertices of the graph after every superstep.
+Further, Bagel also provides a combiner utility where in a reduction operation
+it is applied on the outgoing message which reduces the communication overhead.
+The aggregator utility helps in reducing the number of messages across all
+nodes for all vertices of the graph after every superstep.
 
-\emph{Hash Partitioner}:The algorithm makes use of a hash partitioner which is a Controlled data partitioning utility that reduces communication cost.Also, it keeps the information of how data was partitioned across jobs for further reduction of communication cost.
+\emph{Hash Partitioner}: The algorithm makes use of a hash partitioner which is
+a controlled data partitioning utility that reduces communication cost. Also,
+it keeps the information of how data was partitioned across jobs for further
+reduction of communication cost.
 
-\emph{Kyro serialization}: For the bagel program a special form of serialization technique was used, this  serializes the object sent over the network that will be cached in serialized form.
-The kryo serialization is much faster than the default java serialization used in the spark system as it the communication cost.
+\emph{Kyro serialization}: For the Bagel program a special form of
+serialization technique was used, this serializes the object sent over the
+network that will be cached in serialized form. The Kryo serialization is much
+faster than the default Java serialization used in the Spark system as it adds
+to the communication cost.
 
@@
 Spark installation.\footnote{
-\url{https://groups.google.com/forum/#!topic/spark-users/-nsoF7oonOM}}
+\url{https://groups.google.com/forum/#!topic/Spark-users/-nsoF7oonOM}}
 
@@
         \centering
-        \includegraphics[width=90mm]{images/bagelsimple.png}
+        \includegraphics[width=90mm]{images/Bagelsimple.png}
         \caption{Bagel VS. Simple PageRank}
-        \label{fig:bagelsimple}
+        \label{fig:Bagelsimple}
     \end{figure}
@@
 30 million links in total with sufficient memory. In Figure
-\ref{fig:bagelsimple} regardless of time, we can see that bagel version
+\ref{fig:Bagelsimple} regardless of time, we can see that Bagel version
 outperforms simple version a little bit. The main reason is that for a smaller
@@
 algorithms and degrade linearly with insufficient memory. Using a partitioner,
-the dataset is uniformly distributed across the nodes to avoidance of possible
+the dataset is uniformly distributed across the nodes to avoid the possible
 stragglers. For the scalability test, the time increase with the number of nodes
@@
 solving the akka error would enable the program to run on larger datasets, and
-show great scalability nodes counts of 10 and under.
+show great scalability over nodes counts of 10 and under.
 
@@
 
-% workers and master communication over spark protocol
+% workers and master communication over Spark protocol
 
